### 2021-11-27 eda-h.Rmd
created exploratory analysis file

---
title: "Exploratory Analysis"
author: "Henrique Magalhaes Rio"
output: html_notebook
---

```{r}

set.seed(445)
library(leaps)
library(ISLR)
library(tidyverse) 
library(knitr)
library(glmnet) # package for lasso and ridge regression
library(pls)
library(ISLR)
library(ggpubr)
library(MASS)
library(splines)
library(gam)
library(tree)
library(randomForest)
library(gbm)
library(caret)

```

```{r}
train <- read.csv("train.csv")
```