### 2021-11-30 Bagging Model.Rmd
Adding Oriana's models (LASSO, PLS, Bagging)

---
title: "Bagging Model"
author: "Oriana Meldrum"
date: "11/30/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Bagging model

```{r}
#Set up
#reproducibility
set.seed(445)

library(ISLR)
library(tidyr)
library(dplyr)
library(knitr)
library(tree)
library(randomForest) # random forests & bagging

library(readr)
heart <- read_csv("heart.csv")
View(heart)

#60/40 split
heart_df <- sample(x=nrow(heart), size=.40*nrow(heart))
train <- heart[heart_df,]
test <-  heart[-heart_df,]

```

```{r}

bagging <- randomForest(HeartDisease ~ .,data = train, mtry = 19, ntree = 500) 
predict_bag <- predict(bagging, newdata = test)

mean((predict_bag - test$HeartDisease)^2)

#bagging
```

Test MSE is  0.1052255 for bagging

Now trying to visulize the bagging

```{r}

VI <- data.frame(var=names(heart[,-1]), imp=varImp(bagging))

#sort variable importance descending
VI_plot <- VI[order(VI$Overall, decreasing=TRUE),]

#visualize variable importance with horizontal bar plot
barplot(VI_plot$Overall,
        names.arg=rownames(VI_plot),
        horiz=TRUE,
        col='steelblue',
        xlab='Variable Importance')

```