### 2021-11-30 Lasso Model.Rmd

---
title: "LASSO Model DSCI 445 Project"
author: "Oriana Meldrum"
date: "11/26/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Lasso model: from hw 4

```{r}
#install.packages("glmnet")

#reproducibility
set.seed(445)

library(ISLR)
library(leaps)
library(tidyr)
library(ggplot2)
library(glmnet)
library(caret)
library(pls)

library(readr)
heart <- read_csv("heart.csv")
View(heart)

heart_df <- sample(x=nrow(heart), size=.40*nrow(heart))
train <- heart[heart_df,]
test <-  heart[-heart_df,]

```

```{r}
#60% training, 40% testing

train_matrix <- model.matrix(HeartDisease ~ ., data = train)
test_matrix <- model.matrix(HeartDisease ~ ., data = test)
grid <- 10^seq(10, -2, length=100)

lasso <- glmnet(train_matrix, train$HeartDisease, alpha = 1, lambda = grid, thresh = 1e-12)
plot(lasso)

#Here it breaks
#cv_lasso  <- cv.glmnet(train_matrix, train[, "HeartDisease"], alpha=1)
#plot(cv_ridge)
#best_lambda_lasso <- cv_lasso$lambda.min

#lasso_prediction <- predict(lasso, newx = test_matrix, s = best_lambda_lasso)
#lasso_MSE <- mean((lasso_prediction - test[, "HeartDisease"])^2)

#num_zeros <-cv_lasso$nzero[which.min(cv_lasso$cvm)]

#values to use:
#best_lambda_lasso
#lasso_MSE
#num_zeros

```
Test MSE is  00000000000000000 for LASSO

```{r}
#lasso

x <- model.matrix(HeartDisease ~ . - 1, data = heart)
y <- heart$HeartDisease
lasso_heart <- cv.glmnet(x, y, type.measure = "mse")
plot(lasso_heart)

coef(lasso_heart)

sqrt(lasso_heart$cvm[lasso_heart$lambda == lasso_heart$lambda.1se])

```

```{r}

#cv_lasso <- cv.glmnet(heart, heart$HeartDisease, alpha = 1)
#plot(cv_lasso)
#plot(lasso_fit)

#opt_lambda <-cv_lasso$lambda.min
#opt_lambda

#fit.lasso <- glmnet(lasso_fit, Y, alpha = 1)
#predict(fit.lasso, s = opt_lambda, type = "coefficients")[1:11, ]

```