### 2021-11-30 PLS Model.Rmd

---
title: "PLS Model"
author: "Oriana Meldrum"
date: "11/30/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

#Set up
#reproducibility
set.seed(445)

library(ISLR)
library(leaps)
library(tidyr)
library(ggplot2)
library(glmnet)
library(caret)
library(pls)

library(readr)
heart <- read_csv("heart.csv")
View(heart)

#60/40 split
heart_df <- sample(x=nrow(heart), size=.40*nrow(heart))
train <- heart[heart_df,]
test <-  heart[-heart_df,]

```

Fitting a PLS model on the training set with $M$ chosen using 10-fold CV (on the training set only)

```{r}

pls_fit <- plsr(HeartDisease ~ ., data = train, scale = TRUE, validation = "CV")
validationplot(pls_fit, val.type = "MSEP")

pls_prediction <- predict(pls_fit, test, ncomp = 7)
pls_MSE <- mean((test[, "HeartDisease"] - data.frame(pls_prediction))^2)

m0 <- pcr(HeartDisease~.,data =train, scale = TRUE, validation = "CV")
mse <- MSEP(m0)

# 2. Create a plot of the CV MSE (note root MSE is reported) vs. $M$.
mse <- MSEP(m0)
data.frame(M = mse$comps, mse = t(as.data.frame(mse$val))[, "CV"]) %>%
  ggplot() +
  geom_line(aes(M, mse)) +
  geom_point(aes(M, mse))

m_pls <- mse$comps[which.min(as.data.frame(mse$val)[1,])]

#values to use:
mse
m_pls

```

Test MSE is  0.09841 for PLS at 13 variables