### 2021-12-01 logit.Rmd
Ran logistic model

---
title: "logit"
author: "Henrique Rio"
date: "11/30/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(glmnet)
```

```{r}
train <- read.csv("train.csv")
test <- read.csv("test.csv")

model_glm = glm(HeartDisease ~ ., data = train, family = "binomial")

summary(model_glm)
```

```{r}
predicted2 <- predict(model_glm, newdata = test, "response")

predicted2 <- ifelse(predicted2>0.5,1,0)

confm2<-table(predicted=predicted2,True=test$HeartDisease)
confm2
paste("the overall fraction of correct predictions in the test dataset is :",(confm2[1,1]+confm2[2,2])/sum(confm2))
```