### 2021-12-05 randomforest.Rmd
Random Forest Start

---
title: "RandomForest"
author: "Henrique Rio"
date: "12/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(randomForest)
train<-read.csv("train.csv")
test<-read.csv("test.csv")

```

```{r}
rf_fit <- randomForest(HeartDisease ~ ., data = train, mtry = sqrt(ncol(train) - 1), importance = TRUE, type= "classification")

true <- test$HeartDisease
rf <- predict(rf_fit,test,type="class")

testnmse <- mean((rf-true)^2)

paste("Random forest tree test error",testnmse)
```