### 2021-12-05 randomforest.Rmd
Got the forest working

---
title: "RandomForest"
author: "Henrique Rio"
date: "12/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(randomForest)
train<-read.csv("train.csv")
test<-read.csv("test.csv")

```

```{r}
rf_fit <- randomForest(as.factor(HeartDisease) ~ ., data = train, mtry = sqrt(ncol(train) - 1), importance = TRUE, type= "classification")

true <- test$HeartDisease
rf <- predict(rf_fit,test,type="class")

predicted2 <- ifelse(rf>0.5,1,0)

confm2<-table(predicted=rf,True=true)
confm2

paste("the overall fraction of correct predictions in the test dataset is :",(confm2[1,1]+confm2[2,2])/sum(confm2))
```

```{r}
print(rf_fit)
```

```{r}
plot(rf_fit)
```