### 2021-12-05 randomforest.Rmd
Representative Tree plot

---
title: "RandomForest"
author: "Henrique Rio"
date: "12/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(randomForest)
train<-read.csv("train.csv")
test<-read.csv("test.csv")

options(repos='http://cran.rstudio.org')
have.packages <- installed.packages()
cran.packages <- c('devtools','plotrix','randomForest','tree')
to.install <- setdiff(cran.packages, have.packages[,1])
if(length(to.install)>0) install.packages(to.install)

library(devtools)
if(!('reprtree' %in% installed.packages())){
  install_github('araastat/reprtree')
}
for(p in c(cran.packages, 'reprtree')) eval(substitute(library(pkg), list(pkg=p)))

```

```{r}
library(reprtree)
```

```{r}
rf_fit <- randomForest(as.factor(HeartDisease) ~ ., data = train, mtry = sqrt(ncol(train) - 1), importance = TRUE, type= "classification")

true <- test$HeartDisease
rf <- predict(rf_fit,test,type="class")

predicted2 <- ifelse(rf>0.5,1,0)

confm2<-table(predicted=rf,True=true)
confm2

paste("the overall fraction of correct predictions in the test dataset is :",(confm2[1,1]+confm2[2,2])/sum(confm2))
```

```{r}
print(rf_fit)
```

```{r}
plot(rf_fit)
```

```{r}
reprtree:::plot.getTree(rf_fit)
```

https://stats.stackexchange.com/questions/41443/how-to-actually-plot-a-sample-tree-from-randomforestgettree

Identifying representative trees from ensembles
Banerjee, Mousumi ; Ding, Ying ; Noone, Anne-Michelle
Chichester, UK: John Wiley & Sons, Ltd
Statistics in medicine, 2012-07-10, Vol.31 (15), p.1601-1616