### 2021-12-06 LDA.Rmd
Added the LDA model
My LDA model had a test accuracy of 85.5% and MSE of 0.144. I am worried I did something wrong to get such a low MSE value.

---
title: "LDA_model"
author: "Jorie Alvis"
date: "December 1, 2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)

library(MASS) # for lda

set.seed(445)
```

```{r Data, include=FALSE}
heart_train = read.csv("train.csv")

heart_train$Sex = factor(heart_train$Sex)
heart_train$ChestPainType = factor(heart_train$ChestPainType)
heart_train$FastingBS = factor(heart_train$FastingBS)
heart_train$RestingECG = factor(heart_train$RestingECG)
heart_train$ExerciseAngina = factor(heart_train$ExerciseAngina)
heart_train$ST_Slope = factor(heart_train$ST_Slope)
heart_train$HeartDisease = factor(heart_train$HeartDisease)

heart_test = read.csv("test.csv")

heart_test$Sex = factor(heart_test$Sex)
heart_test$ChestPainType = factor(heart_test$ChestPainType)
heart_test$FastingBS = factor(heart_test$FastingBS)
heart_test$RestingECG = factor(heart_test$RestingECG)
heart_test$ExerciseAngina = factor(heart_test$ExerciseAngina)
heart_test$ST_Slope = factor(heart_test$ST_Slope)
heart_test$HeartDisease = factor(heart_test$HeartDisease)
```

```{r LDA}
LDA_model = lda(HeartDisease~., data=heart_train)

preds = predict(LDA_model, heart_train)
acc = sum(ifelse(preds$class == heart_train$HeartDisease, 1, 0))/nrow(heart_train) # 88.2%

table(preds$class, heart_train$HeartDisease)
acc
```

```{r Testing}
preds = predict(LDA_model, heart_test)
acc = sum(ifelse(preds$class == heart_test$HeartDisease, 1, 0))/nrow(heart_test) # 85.5%

MSE = sum((as.integer(heart_test$HeartDisease)-as.integer(preds$class))^2)/nrow(heart_test) # 0.1444

table(preds$class, heart_test$HeartDisease)
acc
MSE
```

The MSE seems to be way to low for our accuracy. (look at SVM accuracy and MSE)