### 2021-12-06 SVM.Rmd

---
title: "SVM_model"
author: "Jorie Alvis"
date: "December 1, 2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)

library(e1071) # svm library

set.seed(445)
```

```{r Data, include=FALSE}
heart_train = read.csv("train.csv")

heart_train$Sex = factor(heart_train$Sex)
heart_train$ChestPainType = factor(heart_train$ChestPainType)
heart_train$FastingBS = factor(heart_train$FastingBS)
heart_train$RestingECG = factor(heart_train$RestingECG)
heart_train$ExerciseAngina = factor(heart_train$ExerciseAngina)
heart_train$ST_Slope = factor(heart_train$ST_Slope)
heart_train$HeartDisease = factor(heart_train$HeartDisease)
```

```{r SVM_Linear}
# Train a Linear SVM and find the best cost value
# We set scale=TRUE becuase the continuous varibles seem to be normally distrabuted (from EDA)
svm_linear_cv = tune(svm, HeartDisease~., scale=TRUE, kernel="linear", data=heart_train,
                      ranges = list(cost = seq(0.1, 5, by = 0.1)))
svm_linear = svm_linear_cv$best.model

preds = predict(svm_linear, heart_train, type="class")

table(heart_train$HeartDisease, preds)
```

```{r SVM_Polynomial}
# Train a Polynomial SVM
# Find the best cost and degree parameter
svm_poly_cv = tune(svm, HeartDisease~., scale=TRUE, kernel="polynomial", data=heart_train,
                      ranges = list(cost = seq(0.1, 5, by = 0.1),
                                    degree = c(2,3,4, 5)))
svm_poly = svm_poly_cv$best.model

preds = predict(svm_poly, heart_train, type="class")

table(heart_train$HeartDisease, preds)
```

```{r SVM_Radial}
# Train a Radial SVM
# Find the best cost and gamma parameter
svm_radial_cv = tune(svm, HeartDisease~., scale=TRUE, kernel="radial", data=heart_train,
                      ranges = list(cost = seq(0.2, 5, by = 0.2),
                                    gamma = seq(0.25, 3, by = 0.25)))
svm_radial = svm_radial_cv$best.model

preds = predict(svm_radial, heart_train, type="class")

table(heart_train$HeartDisease, preds)
```

```{r Testing}
heart_test = read.csv("test.csv")

preds_liner = predict(svm_linear, heart_test, type="class")
acc_liner = sum(ifelse(preds_liner == heart_test$HeartDisease, 1, 0))/nrow(heart_test)

preds_poly = predict(svm_poly, heart_test, type="class")
acc_poly = sum(ifelse(preds_poly == heart_test$HeartDisease, 1, 0))/nrow(heart_test)

preds_radial = predict(svm_radial, heart_test, type="class")
acc_radial = sum(ifelse(preds_radial == heart_test$HeartDisease, 1, 0))/nrow(heart_test) #87.2%

c("Linear Svm" = acc_liner, "Polynomial Svm" = acc_poly, "Radial SVM" = acc_radial)
```

The SVM kernal with the highest auccuracy was the radial kernal with a `r round(acc_radial, 3)*100`% accuracy. The Radial SVM had a cost parameter of `r svm_radial$cost` and gamma parameter of `r svm_radial$gamma`.

```{r Best_SVM}
table(heart_test$HeartDisease, preds_radial)
radial_mse = sum((heart_test$HeartDisease-as.integer(preds_radial))^2)/nrow(heart_test) #1.253
```

We can see from the confusion matrix of our Radial SVM that there is little bias in the miss-classifications while the MSE is `r round(radial_mse, 4)`.