### 2021-12-06 KNN Model.Rmd

---
title: "KNN Model"
author: "Oriana Meldurm"
date: "12/5/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}

library(ISLR)
library(class)
library(tidyr)

install.packages("fastDummies")
library(fastDummies)

temp_train <- dummy_cols(train, select_columns = c("Sex", "ChestPainType", "RestingECG", "ST_Slope", "ExerciseAngina") )
temp_test <- dummy_cols(test, select_columns = c("Sex", "ChestPainType", "RestingECG", "ST_Slope", "ExerciseAngina"))

drops <- c("Sex", "ChestPainType", "RestingECG", "ST_Slope", "ExerciseAngina")
temp_train <- temp_train[, !(names(temp_train) %in% drops)]
temp_test <-temp_test[, !(names(temp_test) %in% drops)]

trainknn <- temp_train[-temp_train$HeartDisease]
testknn <- test[-test$HeartDisease]

trainknn
testknn

?knn.cv
num_k <- seq(1:100)

my_cv <- knn.cv(temp_train, temp_train$HeartDisease, k=num_k)
my_cv

knn_pred_1 <- knn(temp_train, temp_test, temp_train$HeartDisease, k=11)

confm_knn_1<-table(knn_pred_1,test$HeartDisease)

overall_knn_1 <- (confm_knn[1,1]+confm_knn[2,2])/sum(confm_knn)

overall_knn_1

 #confusion matrix
kable(confm_knn)
mean(knn_pred !=test$HeartDisease)

```

```{r}

library(caret)

#5 fold
trControl <- trainControl(method = "cv", number = 5)

fit <- train(as.factor(HeartDisease) ~.,
             method = "knn", 
             tuneGrid = expand.grid(k = 1:100), 
             trControl = trControl, 
             metric = "Accuracy", 
             data = temp_train)

fit
```