### 2021-12-07 boosting.Rmd
boosting file

---
title: "Boosting"
author: "Henrique"
output: html_notebook
---

```{r}

train <- read.csv("train.csv")
test <- read.csv("test.csv")
library(gbm)
library(caret)

```

```{r}

train$ChestPainType <- as.factor(train$ChestPainType)
train$Sex <- as.factor(train$Sex)
train$FastingBS <- as.factor(train$FastingBS)
train$RestingECG <- as.factor(train$RestingECG)
train$ExerciseAngina <- as.factor(train$ExerciseAngina)
train$ST_Slope <- as.factor(train$ST_Slope)

test$ChestPainType <- as.factor(test$ChestPainType)
test$Sex <- as.factor(test$Sex)
test$FastingBS <- as.factor(test$FastingBS)
test$RestingECG <- as.factor(test$RestingECG)
test$ExerciseAngina <- as.factor(test$ExerciseAngina)
test$ST_Slope <- as.factor(test$ST_Slope)

set.seed(445)
lambda=seq(from=0,to=1,length.out=100)

tmse<-0

boost <- gbm(HeartDisease~.,data=temp_train,n.trees=1000,shrinkage=4,interaction.depth = 2 ,distribution = "bernoulli")

gbm_grid =  expand.grid(interaction.depth = 1:5,
                        n.trees = (1:6) * 500,
                        shrinkage = c(0.001, 0.01, 0.1),
                        n.minobsinnode = 10)

```