### 2021-12-07 EDA-J.Rmd
Created EDA markdown
Created an R markdown for EDA for our training dataset

---
title: "EDA-J"
author: "Jorie Alvis"
date: "December 7, 2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)

set.seed(445)

heart_train = read.csv("train.csv")
```

```{r Cor}
pairs(heart_train)
pairs(heart_train[,-c(2,3,6,7,9,10,11)])
```

```{r Norm, echo=FALSE}
qqnorm(heart_train$Age, main="QQ Plot of Age")
qqline(heart_train$Age)

qqnorm(heart_train$RestingBP, main="QQ Plot of Resting BP")
qqline(heart_train$RestingBP)

qqnorm(heart_train$Cholesterol, main="QQ Plot of Cholesterol")
qqline(heart_train$Cholesterol)

qqnorm(heart_train$MaxHR, main="QQ Plot of Max HR")
qqline(heart_train$MaxHR)
```