### 2021-12-07 Boosting-J.Rmd
Renamed Jorie's boosting model file
Renamed my boosting model R file to not conflict with Henrique's

---
title: "Boosting"
author: "Jorie Alvis"
date: "December 5, 2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)

library(gbm) # boosting package

set.seed(445)
```

```{r Data, include=FALSE}
heart_train = read.csv("train.csv")

heart_train$Sex = factor(heart_train$Sex)
heart_train$ChestPainType = factor(heart_train$ChestPainType)
heart_train$FastingBS = factor(heart_train$FastingBS)
heart_train$RestingECG = factor(heart_train$RestingECG)
heart_train$ExerciseAngina = factor(heart_train$ExerciseAngina)
heart_train$ST_Slope = factor(heart_train$ST_Slope)
#heart_train$HeartDisease = factor(heart_train$HeartDisease)
heart_train$HeartDisease = as.integer(heart_train$HeartDisease)

heart_test = read.csv("test.csv")

heart_test$Sex = factor(heart_test$Sex)
heart_test$ChestPainType = factor(heart_test$ChestPainType)
heart_test$FastingBS = factor(heart_test$FastingBS)
heart_test$RestingECG = factor(heart_test$RestingECG)
heart_test$ExerciseAngina = factor(heart_test$ExerciseAngina)
heart_test$ST_Slope = factor(heart_test$ST_Slope)
#heart_test$HeartDisease = factor(heart_test$HeartDisease)
heart_test$HeartDisease = as.integer(heart_test$HeartDisease)
```

```{r Boosting}
# The response needs to be an integer not a factor
boosted_model = gbm(HeartDisease~., data=heart_train, n.trees=10000, interaction.depth=3,
                    shrinkage=0.01, distribution="bernoulli", cv.folds=10)
gbm.perf(boosted_model, plot.it = TRUE, method="cv")
```

```{r Testing}
t=gbm(HeartDisease~., data=heart_train, n.trees=1000, interaction.depth=3,
                    shrinkage=0.075, distribution="bernoulli")

preds = predict(t, heart_test, type="response")
preds2 = ifelse(preds<0.5, 0, 1)

acc = sum(as.integer(preds2==heart_test$HeartDisease))/nrow(heart_test)

table(preds2, "True"=heart_test$HeartDisease)
acc
summary(t)
```