### 2021-12-07 presentation.Rmd
CHANGES

---
title: "Group 9 Presentation"
author: "Henrique, Oriana, Jorie"
date: "12/4/2021"
output:
  slidy_presentation: default
  ioslides_presentation: default
  beamer_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
options(repos='http://cran.rstudio.org')
have.packages <- installed.packages()
cran.packages <- c('devtools','plotrix','randomForest','tree')
to.install <- setdiff(cran.packages, have.packages[,1])
if(length(to.install)>0) install.packages(to.install)

library(devtools)
if(!('reprtree' %in% installed.packages())){
  install_github('araastat/reprtree')
}
for(p in c(cran.packages, 'reprtree')) eval(substitute(library(pkg), list(pkg=p)))
```

```{r, include=FALSE}

#reproducibility
set.seed(445)

library(ISLR)
library(leaps)
library(tidyr)
library(dplyr)
library(ggplot2)
library(glmnet)
library(caret)
library(pls) 
library(tree)
library(randomForest)
library(knitr)
library(class)
library(fastDummies)
library(reprtree)

train <- read.csv("train.csv")
test <- read.csv("test.csv")

train$ChestPainType <- as.factor(train$ChestPainType)
train$Sex <- as.factor(train$Sex)
train$FastingBS <- as.factor(train$FastingBS)
train$RestingECG <- as.factor(train$RestingECG)
train$ExerciseAngina <- as.factor(train$ExerciseAngina)
train$ST_Slope <- as.factor(train$ST_Slope)
train$HeartDisease <- as.factor(train$HeartDisease)

test$ChestPainType <- as.factor(test$ChestPainType)
test$Sex <- as.factor(test$Sex)
test$FastingBS <- as.factor(test$FastingBS)
test$RestingECG <- as.factor(test$RestingECG)
test$ExerciseAngina <- as.factor(test$ExerciseAngina)
test$ST_Slope <- as.factor(test$ST_Slope)
test$HeartDisease <- as.factor(test$HeartDisease)

```

## Group 9

Heart Disease
Oriana Meldrum, Henrique Rio, Jorie Alvis

## Introduction

In 2019 659,041 Americans died of heart disease. Four out of five heart disease deaths are due to heart attacks and strokes, and one-third of these deaths occur prematurely in people under 70 years of age. Heart failure is a common event caused by heart disease and this dataset we used contains 11 variables that can be used to predict a possible heart disease.

People with cardiovascular disease or who are at high cardiovascular risk (due to the presence of one or more risk factors) need early detection and management so machine learning techniques are of great clinical intrest for early detection.

(https://www.cdc.gov/nchs/fastats/leading-causes-of-death.htm, https://www.kaggle.com/fedesoriano/heart-failure-prediction)

## Data {.smaller}

- There are 918 observations which we split into 60/40 for training and validation.

- 12 predictors are included in the data which are:

+ Age

+ Sex M/F

+ Chest Pain Type:

++ TA: Typical Angina

++ ATA: Atypical Angina

++ NAP: Non-Anginal Pain

++ ASY: Asymptomatic

+ Resting Blood Pressure [mm Hg]

+ Cholesterol [mm/dl]

+ Fasting Blood Sugar Y/N

## Data {.smaller}

+ Resting Electrocardiogram

++ Normal: Normal

++ ST: Having ST-T Wave abnormality

++ LVH: Showing probable or definite left ventricular hypertrophy.

+ Max HR: Maximum heart rate achieved.

+ Exercise Angina: exercise-induced angine Y/N

+ Old peak: Oldpeak= ST value of depression

+ ST_Slope:

++ Up sloping(up)

++ Flat(flat)

++ Down sloping (down)

+ Heart Disease: [1:Heart Disease, 0: Normal]

## LDA

Jorie's LDA

## Random Forest

```{r}
rf_fit <- randomForest(as.factor(HeartDisease) ~ ., data = train, mtry = sqrt(ncol(train)-1), importance = TRUE, type= "classification")

true <- test$HeartDisease
rf <- predict(rf_fit,test,type="class")

predicted2 <- ifelse(rf>0.5,1,0)

confm2<-table(predicted=rf,True=true)
kable(confm2)

paste("the overall fraction of correct predictions in the test dataset is :",(confm2[1,1]+confm2[2,2])/sum(confm2))

rf_fit

imp<-data.frame(rf_fit$importance)

imp$variables <- rownames(imp)
rownames(imp) <- 1:nrow(imp)

imp <- imp[order(imp$MeanDecreaseAccuracy),]

```

## Random Forest

- Variable importance shows.

```{r}
ggplot(imp)+geom_col(aes(x=variables,y=MeanDecreaseAccuracy,fill=variables))+coord_flip()+labs(title="Variable Importance",y=" Mean decrease in Accuracy", x= " Variable")+scale_y_continuous(labels = scales::percent_format())
```

## Representative Tree
- Following the paper by  Banerjee, et al.(2011) and the package reprtree we can get a representative tree from the random forest.
- This tree is calculated using a measure of similarity in which 2 trees are similar if they have the same prediction for all observations.
- Using this idea on the random forest we get the following tree:

## Tree
```{r fig.height=7, fig.width=15, message=FALSE, warning=FALSE,echo=FALSE}
reptree <- ReprTree(rf_fit, train, metric='d2')

plot(reptree, index=1)

```

## SVM

Jorie's SVM

## (the new thing)

Henrique's new thing insted of PCA

## Logistic Regression

Henrique's Logisic Regression

```{r}
model_glm = glm(HeartDisease ~ ., data = train, family = "binomial")
#summary(model_glm)

predicted2 <- predict(model_glm, newdata = test, "response")
predicted2 <- ifelse(predicted2>0.5,1,0)
confm2<-table(predicted=predicted2,True=test$HeartDisease)
kable(confm2)

paste("The overall fraction of correct predictions in the test dataset is :",(confm2[1,1]+confm2[2,2])/sum(confm2))

 #MSE

```

## Boosting

Jories's Boosting

## LASSO

Oriana's Lasso

```{r}

train_matrix <- model.matrix(HeartDisease ~ ., data = train)
test_matrix <- model.matrix(HeartDisease ~ ., data = test)
grid <- 10^seq(10, -2, length=100)

lasso <- glmnet(train_matrix, train$HeartDisease, alpha = 1, lambda = grid, thresh = 1e-12, family = "binomial")
#plot(lasso)

cv_lasso  <- cv.glmnet(train_matrix, train[, "HeartDisease"], alpha=1,family = "binomial")
#plot(cv_lasso)
best_lambda_lasso <- cv_lasso$lambda.min

lasso_prediction <- predict(lasso, newx = test_matrix, s = best_lambda_lasso)
lasso_MSE <- mean((lasso_prediction - test[, "HeartDisease"])^2)

num_zeros <-cv_lasso$nzero[which.min(cv_lasso$cvm)]

# MSE values to use:
best_lambda_lasso
lasso_MSE
num_zeros

#Confusion Matrix
predicted_lasso <- predict(lasso, newx= test_matrix, s=best_lambda_lasso)
predicted_lasso <- ifelse(predicted_lasso>0.5,1,0)
confm_lasso<-table(predicted=predicted_lasso,True=test$HeartDisease)
confm_lasso

#Over all correct:
overall_lasso <- (confm_lasso[1,1]+confm_lasso[2,2])/sum(confm_lasso)

kable(confm_lasso) #confusion matrix
overall_lasso #over all correct
```

## KNN

```{r}

temp_train <- dummy_cols(train, select_columns = c("Sex", "ChestPainType", "RestingECG", "ST_Slope", "ExerciseAngina") )
temp_test <- dummy_cols(test, select_columns = c("Sex", "ChestPainType", "RestingECG", "ST_Slope", "ExerciseAngina"))

drops <- c("Sex", "ChestPainType", "RestingECG", "ST_Slope", "ExerciseAngina")
temp_train <- temp_train[, !(names(temp_train) %in% drops)]
temp_test <-temp_test[, !(names(temp_test) %in% drops)]

#trainknn <- temp_train[-temp_train$HeartDisease]
#testknn <- test[-test$HeartDisease]

num_k <- seq(1:100)

my_cv <- knn.cv(temp_train, temp_train$HeartDisease, k=num_k)

knn_pred_1 <- knn(temp_train, temp_test, temp_train$HeartDisease, k=11)

confm_knn_1<-table(knn_pred_1,test$HeartDisease)

overall_knn_1 <- (confm_knn_1[1,1]+confm_knn_1[2,2])/sum(confm_knn_1)

overall_knn_1

 #confusion matrix
kable(confm_knn_1)
mean(knn_pred_1 !=test$HeartDisease)
```

## Bagging
```{r,echo=FALSE,warning=FALSE,fig.width=6,fig.height=3}

bagging <- randomForest(HeartDisease ~ .,data = train, mtry = 19, ntree = 500, type = "class", 
                        importance = TRUE,
                        proximity = TRUE)

#Confusion Matrix
predict_bag <- predict(bagging, newdata = test)
predicted_bag <- ifelse(predict_bag>0.5,1,0)
confm_bag<-table(predict_bag,test$HeartDisease)
confm_bag

#MSE
bagging_mse <- mean((predict_bag - test$HeartDisease)^2)

#Over all correct:
overall_bag <- (confm_bag[1,1]+confm_bag[2,2])/sum(confm_bag)

kable(confm_bag) #confusion matrix
overall_bag #over all correct

#Visulizaton of bagging
VI <- data.frame(var=names(test[,-12]), imp=varImp(bagging))

#sort variable importance descending
VI_plot <- VI[order(VI$imp.0, decreasing=TRUE),]

#visualize variable importance with horizontal bar plot
#barplot(VI_plot$imp.0,
   #     names.arg=rownames(VI_plot),
    #    horiz=TRUE,
     #   col='steelblue',
      #  xlab='Variable Importance')

ggplot(VI_plot)+geom_col(aes(x=(rownames(VI_plot)),y=imp.0,fill=rownames(VI_plot)))+coord_flip()+guides(fill=FALSE)+
  labs(x="Variable Name",y="Importance")

```

## Comparision of all method's MSE

Chart with all MSE

```{r}
library(knitr)
df1 = data.frame(Method = c('LDA', 'RF', 'SVM', 'new THING', 'Logistic Regression', 'Boosting', 'LASSO', 'PLS', 'Bagging'),  MSE = c(0,0,0,0,0.1498638,0,0,0.1054 ,0.1165492))

kable(df1)

```

## Conclusion

the METHOD worked best, with an MSE of SOMETHING